---
title: "eCollection Analysis"
author: "MOF DS Team"
date: "10/15/2019"
output:
  html_document:
    theme: flatly
    toc: yes
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = FALSE,
    message = FALSE,
    warning = FALSE,
    dpi = 300)
```


```{r, libraries}
# Load Package and Libraries

# Library for data manipulation and Visualization
library(tidyverse)
library(gmodels)
library(forcats)
library(purrr)
library(lubridate)
library(tidyquant)

# Library for reading and loading data
library(readr)
library(readxl)
library(writexl)
library(stringr)
library(stringi)
library(fs)

# Library for Machine Learning Packages
library(parsnip)
library(class)
library(rpart)
library(rpart.plot)
library(ggrepel)
library(broom)
library(umap)
library(rsample)
library(yardstick)
library(gmodels)
library(tinytex)
#tinytex::install_tinytex()


# ecol_pymt_tbl <- read_excel("00_Data/ecollection_payments.xlsx", sheet = 1)
ecol_pymt_tbl <- read_rds(path = "00_Data/ecol_pymt_tbl.rds")
umap_kmeans_3_result_tbl <- read_rds("00_Data/umap_kmeans_3_result_tbl.rds")
ecol_class <- read_rds("00_Data/ecol_class.rds")
```

### Knowledge, Findings and Highlights

* **Internet and ATM** Channels Accounts for the Largest None Delivered Payment
* **Ministry of Mineral Resources** Accounts for the Largest None Delivered Payment Amount
* **Ministry of Finance** Have the Highest Payments Amount for a Ten (10) years Period (2009 - 2019)
* In **2018** Ministry of Finance **Declined** in **Payments Amount** and The Ministry of Mineral Resources **Increased** in Payment Amount in **2018 and 2019**
* **Gouf Province Municipal** Have the **Highest Number of Payments** but **low Payment Amount** than other Entities like **MOF and Ministry of Mineral Resources**
* **Internet and ATM Channel** __Grew__ for Payments Above SAR 500K from **2018 to 2019**
* Every **3 to 4 years** there seems to be a **fluctuation** in Payment Amount (Increase or Decrease) Across Channels
* **Reconciled Payments** Is the **Largest** Across Payment Status
* **Bank Teller Channel** is the Top Performing Channel with the **largest** Payment Amount
* **Gouf Province Municipal** Accounts for the largest number of payment made with **Electronic Channels** (ATM, Internet, Mobile)
* **Large Payments** Get's Delivered **Immediately** and most **Smaller Payments** takes longer time to get **delivered**
* **Delayed Payments** are **Prevalent** with **POS, ATM and Internet** Channels
* **Ministry of Finance and Gouf Province Municipal** have the largest number of **delayed** payment

* Payments Between **SAR 500 and SAR 1000** should be done with **CAM, PDA and POS**
* It is not By Chance that **Ministry of Mineral Resources** took the lead with the highest payment amount in **2018 and 2019** from **Ministry of Finance**. Therefore further investigation is required to know what really happened
* Delivery time for **smaller payment** should be **Immediately** 





```{r, transform data}
ecol_pay_tbl <- ecol_pymt_tbl %>% 
    select(-c(eclctn_pmt_ref_num, 
              account_id, 
              pmtmthd_trn_id, 
              pmtmthd_bnk_trn_num, 
              agency_id, account_id, settl_dt, rvsrl_dt, 
              pmt_ref_info, dlvry_req_id, dlvry_ts, 
              dlvry_error_code, dlvry_retry_count, 
              calc_fees, calc_vat_amt, vat_number, last_batch_id, 
              last_job_run_id, job_id, migrationdate, 
              bill_cycle, cust_id, cust_id_type, recon_req_id, abc))

ecol_pay_tbl <- ecol_pay_tbl %>% 
                select(-c(recon_date_hijri, ismigrated, 
                          as_of_date, bill_number, send_dlvry_ts,
                          recon_dt))


ecol_pay_tbl <- ecol_pay_tbl %>% 
    set_names(. %>% str_to_title()) %>% 
    mutate(Amount = as.numeric(Cur_amt)) %>% 
    select(-Cur_amt) %>% 
    mutate(Partner_code = as_factor(Partner_code)) %>% 
    mutate(Pmtmthd_bnk_id = as_factor(Pmtmthd_bnk_id)) %>% 
    mutate(Pmtmthd_pmt_channel = as_factor(Pmtmthd_pmt_channel)) %>% 
    mutate(Pmt_dt = ymd_hms(Pmt_dt) %>% as.Date()) %>% 
    mutate(Source_Channel = as_factor(Source_channel)) %>% 
    mutate(Pmt_method = as_factor(Pmt_method)) %>% 
    mutate(Dlvry_status = as_factor(Dlvry_status)) %>% 
    mutate(Last_update_date = ymd_hms(Last_update_date)) %>% 
    mutate(State = as_factor(State)) %>% 
    mutate(Creation_date = ymd_hms(Creation_date)) %>% 
    mutate(Send_flag = as_factor(Send_flag)) %>% 
    mutate(Send_dlvry_state = as_factor(Send_dlvry_state))

ecol_pay_tbl <- ecol_pay_tbl %>% 
    mutate(Pmtmthd_pmt_channel = case_when(
        Pmtmthd_pmt_channel == "BTELLER" ~ "Bank Teller",
        Pmtmthd_pmt_channel == "IVR" ~ "Interactive Voice Response",
        Pmtmthd_pmt_channel == "CAM" ~ "Common Area Maintenance",
        Pmtmthd_pmt_channel == "CORP" ~ "Corporate Payment",
        Pmtmthd_pmt_channel == "PDA" ~ "Payment Distribution Agencies",
        Pmtmthd_pmt_channel == "POS" ~ "Point of Sales",
        TRUE ~ as.character(Pmtmthd_pmt_channel)
    ))


ecol_pay_tbl <- ecol_pay_tbl %>% 
    mutate(Pmtmthd_pmt_channel = as_factor(Pmtmthd_pmt_channel))


ecol_pay_tbl <- ecol_pay_tbl %>% 
    mutate(Partner_code = case_when(
        Partner_code == "DMMR" ~ "Ministry of Mineral Resources",
        Partner_code == "GEPP" ~ "eProcurement Portal",
        Partner_code == "GPM" ~ "Gouf Province Municipal",
        Partner_code == "MOE" ~ "Ministry of Education",
        Partner_code == "MOF" ~ "Ministry of Finance",
        Partner_code == "NBPM" ~ "Northern Boarder Municipal",
        Partner_code == "PRC" ~ "Premium Residency Center",
        Partner_code == "PTA" ~ "Public Transport Agency",
        Partner_code == "REGA" ~ "Real Estate Government Agency",
        Partner_code == "SAIP" ~ "Saudi Intellectual Property",
        Partner_code == "SEDA" ~ "Saudi Export Development Agency",
        Partner_code == "SIO" ~ "Saudi Irrigation Organization"
    ))


ecol_pay_tbl <- ecol_pay_tbl %>% 
  select(-Main_agency_id) %>% 
  mutate(Send_flag = case_when(
    Send_flag == "Y" ~ "Yes",
    Send_flag == "N" ~ "No"
  )) %>%
  # Removing some date columns we don't need
  
  # select(-Settl_dt, Dlvry_ts, -Send_dlvry_ts) %>% 
  
  # Creating a duration column from the creation date and last updated date for payment
  
mutate(duration = round(interval(Creation_date, Last_update_date)/ddays(1), digits = 2)) %>%
mutate(duration_minutes = round
       (interval(Creation_date, Last_update_date)/dminutes(1), digits = 2))

```






### None Delivered Payments Across Channels
**Internet Channel** have the **Highest Payment Amount** not delivered than other channels
```{r, bar plot}
ecol_pay_tbl %>%
    filter(Dlvry_status == "Not Delivered") %>% 
    filter(!is.na(Pmtmthd_pmt_channel)) %>% 
    group_by(Pmtmthd_pmt_channel, Partner_code) %>% 
    summarise(Amount = sum(Amount)) %>% 
    ungroup() %>% 
    ggplot(aes(Pmtmthd_pmt_channel, Amount)) +
    geom_col() +
    theme_tq() +
    scale_fill_tq() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(x = "",
         y = "Payment Amount",
         title = "None Delivered Payments Across Channels"
         )

```









### Entity with High None Delivered Payments
**Ministry of Mineral Resources** Have the highest **None Delivered Payment** than other **Entities**
```{r}
ecol_pay_tbl %>%
    filter(Dlvry_status == "Not Delivered") %>% 
    filter(!is.na(Pmtmthd_pmt_channel)) %>% 
    mutate(Pmtmthd_pmt_channel = case_when(
      Pmtmthd_pmt_channel == "Corporate Payment" ~ "CORP",
      Pmtmthd_pmt_channel == "Interactive Voice Response" ~ "IVR",
      Pmtmthd_pmt_channel == "Common Area Maintenance" ~ "CAM",
      TRUE ~ as.character(Pmtmthd_pmt_channel)
    )) %>% 
    group_by(Pmtmthd_pmt_channel, Partner_code) %>% 
    summarise(Amount = sum(Amount)) %>% 
    ungroup() %>% 
    ggplot(aes(Pmtmthd_pmt_channel, Amount)) +
    geom_col() +
    #coord_flip() +
    facet_wrap(~Partner_code, ncol = 3) +
    theme_tq() +
    scale_fill_tq() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(x = "Payment Channel",
         y = "Payment Amount",
         title = "None Delivered Payments Across Entities"
    )

```








### Volume of None Delivered Payments
The **Volume** of None Delivered Payments is **very High** for **Internet and ATM Payments**
```{r}
ecol_pay_tbl %>%
    filter(Dlvry_status == "Not Delivered") %>% 
    filter(!is.na(Pmtmthd_pmt_channel)) %>% 
    count(Pmtmthd_pmt_channel, Partner_code) %>% 
    ggplot(aes(Pmtmthd_pmt_channel, n)) +
    geom_col() +
    theme_tq() +
    scale_fill_tq() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(x = "",
         y = "Payment Amount",
         title = "Number of None Delivered Payments Across Channels"
    )

```


### Channel Performance
**None Delivered Payments**  are __Prevalent__ with **Internet, BTELLER and ATM Channels**.
**Payment** amounts between **SAR 500 and SAR 1000** are not **Delivered Frequently** with **Internet, BTELLER, ATM and IVR Channels**. 
**CAM, PDA and POS Channels** are more likely suitable for Payments btw SAR 500 and SAR 1000 as they have minimal **None Delivery** payments.

```{r}
ecol_pay_tbl %>% 
    filter(!is.na(Pmtmthd_pmt_channel)) %>% 
    ggplot(aes(Pmtmthd_pmt_channel, Amount)) +
    geom_jitter(width = 0.1, alpha = 0.1) +
    geom_violin(width = 0.8, alpha = 0.5) +
    coord_flip() +
    #scale_y_continuous(labels = scales::comma) +
    scale_y_log10(labels = scales::comma) +
    theme_tq() +
    scale_fill_tq() +
    facet_wrap(~Dlvry_status) +
    labs(x = "Payment Amount",
          y = "Payment Channel",
          title = "POS, ATM, Internet & IVR for payments less than 1 Halal")
```











### Payments Above SAR 10M
Payments above **SAR 10M** are done with **BTELLER and CORP Channels** 
```{r, Boxplt}
ecol_pay_tbl %>% 
    filter(!is.na(Pmtmthd_pmt_channel)) %>% 
    ggplot(aes(Pmtmthd_pmt_channel, Amount)) +
    geom_boxplot() +
    coord_flip() +
    #scale_y_log10(labels = scales::comma)
    scale_y_continuous(labels = scales::comma) +
    labs(x = "Payment Amount",
         y = "Payment Channels",
         title = "BTELLER, CORP, INTERNET and ATM Contains Outliers")
```












### 10 Years Overall Payment Performance
**Ministry of Finance** have the **Highest Payment** amount for a **10 years period**
(2009 - 2019).
```{r}
ecol_pay_tbl %>% 
    ggplot(aes(Partner_code, Amount)) +
    geom_col() +
    scale_y_continuous(labels = scales::comma) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(x = "Entity",
         y = "Payment Amount",
         title = "Ministry of Finance Leads")
```









### Change in Payment Performance
**Ministry of Mineral Resources** have the **Highest Payment** amount in **2019** compared to the **Previous Years**.
**Ministry of Finance** shows a **declining trend** from **2018**
```{r}
ecol_pay_tbl %>% 
    mutate(Year = year(Pmt_dt)) %>% 
    filter(Year %in% c("2016","2017","2018", "2019")) %>%
    group_by(Partner_code, Year) %>% 
    summarise(Amount = sum(Amount)) %>% 
    ggplot(aes(Partner_code, Amount)) +
    geom_col() +
    facet_wrap(~Year) +
    scale_y_continuous(labels = scales::comma) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(x = "Entity",
         y = "Payment Amount",
         title = "MOF Declined in 2018 and MOE Increased in 2019")
```










### 4 Years Volume of Payment for Entities
**Gouf Province Municipal** have the **Largest** volume of payments but **low payment amounts**.
There is a **decrease** in the **Vumber of Payments** for **MOF** which begins in **2018**.
**Ministry of Mineral Reources** shows **growth** in **2018 and 2019** with respect to the **volume of payments**.
```{r}
ecol_pay_tbl %>% 
    mutate(Year = year(Pmt_dt)) %>% 
    filter(Year %in% c("2016","2017","2018", "2019")) %>%
    count(Partner_code, Year) %>% 
    #summarise(Amount = sum(Amount)) %>% 
    ggplot(aes(Partner_code, n)) +
    geom_col() +
    facet_wrap(~Year) +
    #scale_y_continuous(labels = scales::comma) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(x = "Entity",
         y = "Volume of Payments",
         title = "Gouf Province Municipal have the highest number of ePayments")

```









```{r}
ecol_pay_tbl %>% 
  ggplot(aes(duration, Amount)) +
  geom_point() +
  scale_y_continuous(labels = scales::comma) +
  theme_tq() + # making the background have a grid look
  scale_fill_tq() +
  labs(x = "Duration (Days)",
       y = "Payment Amount",
       title = "Large Payments seems to get completed Immediately",
       subtitle = "small payments likely don't get completed Immedately")
```









```{r}
ecol_pay_tbl %>% 
  filter(duration >= 180) %>% 
  filter(!is.na(Pmtmthd_pmt_channel)) %>% 
  count(Pmtmthd_pmt_channel) %>% 
  ggplot(aes(Pmtmthd_pmt_channel, n)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(y = "Number of Delay",
       x = "Channel",
       title = "Delayed Payment Above 6 Months Across Channels")
```










```{r}
ecol_pay_tbl %>% 
  filter(duration >= 180) %>% 
  filter(!is.na(Pmtmthd_pmt_channel)) %>% 
  count(Partner_code) %>% 
  ggplot(aes(Partner_code, n)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Entity",
       y = "Number of Delay",
       title = "Entity with Delayed Payments Above 6 Months")
```









```{r}
ecol_pay_tbl %>% 
  filter(duration >= 90) %>% 
  filter(!is.na(Pmtmthd_pmt_channel)) %>% 
  mutate(Year = year(Creation_date)) %>% 
  count(Pmtmthd_pmt_channel, Year) %>% 
  ggplot(aes(Pmtmthd_pmt_channel, n)) +
  geom_col() +
  facet_wrap(~Year) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Channel",
       y = "Number of Delay",
       title = "Delayed Payments of 3 Months from 2009 to 2019")
```









Low < 1 day, Medium < 26 days, High > 26 days
```{r}
ecol_class %>% 
  filter(!is.na(Pmtmthd_pmt_channel)) %>% 
  count(duration_range, Pmtmthd_pmt_channel) %>% 
  ggplot(aes(duration_range, n, fill = Pmtmthd_pmt_channel)) +
  geom_col(position = "dodge") +
  scale_y_continuous(labels = scales::comma) +
  theme_tq() +
  scale_fill_tq() +
  labs(x = "Duration Range",
       y = "Number of Delays",
       title = "Delayed Payment by Channels")

```











Low < 1 day, Medium < 26 days, High > 26 days
```{r}
ecol_class %>% 
  bind_cols(ecol_pay_tbl %>% select(Creation_date)) %>%
  mutate(Year = year(Creation_date)) %>%
  select(-Creation_date) %>%
  filter(!is.na(Pmtmthd_pmt_channel)) %>% 
  count(duration_range, Year, Pmtmthd_pmt_channel) %>% 
  mutate(duration_range = fct_reorder2(duration_range, n, Year)) %>% 
  ggplot(aes(duration_range, n, fill = Pmtmthd_pmt_channel)) +
  geom_col(position = "dodge") +
  facet_wrap(~Year) +
  scale_y_continuous(labels = scales::comma) +
  theme_tq() +
  scale_fill_tq() +
  labs(x = "Duration Range",
       y = "Number of Delay",
       title = "Delayed Payment Across Years (2009 - 2019)")
```












```{r}
ecol_class %>% 
  filter(!is.na(Pmtmthd_pmt_channel)) %>% 
  count(duration_range, Source_channel) %>% 
  ggplot(aes(duration_range, n, fill = Source_channel)) +
  geom_col(position = "dodge") +
  scale_y_continuous(labels = scales::comma) +
  theme_tq() +
  scale_fill_tq() +
  labs(x = "Duration",
       y = "Number of Delay",
       title = "Delayed Payment by Source Channel")
```










```{r}
ecol_pay_tbl %>% 
    mutate(Pyt_month = month(Pmt_dt, label = TRUE, abbr = TRUE)) %>% 
    group_by(Pyt_month, State) %>% 
    summarise(Amount = sum(Amount)) %>% 
    ggplot(aes(Pyt_month, Amount)) +
    geom_col() +
    facet_wrap(~State) +
    scale_y_continuous(labels = scales::comma) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(x = "Month",
         y = "Payment Amount",
         title = "Large Reconciliation Payments is Happening alot in July")
```











### Payments Above SAR 500K
Payments **BTW SAR 500K and SAR 1M** are done Mostly with **ATM, BTELLER, INTERNET and CORP Channels**. However ATM Channel has more **variable payments** within **SAR 1M**.
```{r, Above 500k}
above_500k <- ecol_pay_tbl %>% 
    filter(Amount > 500000)

    above_500k %>% 
    # mutate(Year = year(Pmt_dt)) %>%
    # filter(Amount >= 1000000) %>% 
    # filter(Year == 2018 | Year == 2019) %>%
    filter(!is.na(Pmtmthd_pmt_channel)) %>% 
    ggplot(aes(Pmtmthd_pmt_channel, Amount)) +
    geom_jitter(width = 0.1) +
    geom_violin(width = 0.8, alpha = 0.5) +
    coord_flip() +
    #scale_y_continuous(labels = scales::comma) +
    scale_y_log10(labels = scales::comma) +
    theme_tq() +
    scale_fill_tq() +
    #facet_wrap(~Year)
    labs(x = "Payment Channel",
         y = "Payment Amount",
         title = "Channel Payments Above SAR 500K")
```










### Payment Preference 2018 & 2019
**ATM and Internet** Payment Channel for Payments above **SAR 500K** __grew__ in **2019** as more payments **above SAR 500K** was done with **ATM and Internet Channels** than in 2018.
**BTELLER** channel remains the Payment Channel of Choice for **large Payments**
```{r}
above_500k %>% 
    mutate(Year = year(Pmt_dt)) %>%
    filter(Year == 2018 | Year == 2019) %>%
    filter(!is.na(Pmtmthd_pmt_channel)) %>% 
    ggplot(aes(Pmtmthd_pmt_channel, Amount)) +
    geom_jitter(width = 0.1) +
    geom_violin(width = 0.8, alpha = 0.5) +
    coord_flip() +
    #scale_y_continuous(labels = scales::comma) +
    scale_y_log10(labels = scales::comma) +
    theme_tq() +
    scale_fill_tq() +
    facet_wrap(~Year) +
    labs(x = "Payment Channel",
         y = "Payment Amount",
         title = "SAR 500K Payment Preference Growth BTW 2018 and 2019")
```










### Payment Channel Trends
There is likely a **3 - 4 Years fluctuating trends** across **Top Channels**, Therefore Payment Amount **fluctuations** **(upward or downward)** are most like to happen every **3 - 4 years for Top Channels (CORP, POS, INTERNET, BTELLER, ATM and IVR)**.
**CORP and Internet Channel** are on the **upward** trend while other channels are currently on the **downward trend**.
```{r}
ecol_pay_tbl %>% 
    filter(!is.na(Pmtmthd_pmt_channel)) %>% 
    mutate(pymt_date = floor_date(Pmt_dt, "year") %>% ymd()) %>% 
    group_by(pymt_date, Pmtmthd_pmt_channel) %>% 
    summarise(Amount = sum(Amount)) %>% 
    
    ungroup() %>% 
    
    ggplot(aes(pymt_date, Amount)) +
    geom_point() +
    geom_line(size = 0.5, linetype = 1) +
    geom_smooth(span = 0.2, method = "loess") +
    facet_wrap(~Pmtmthd_pmt_channel, ncol = 3) +
    expand_limits(y = 0) +
    scale_y_continuous(labels = scales::comma) +
    theme_tq() +
    scale_fill_tq() +
    labs(x = "Year",
         y = "Payment Amount",
         title = "Payment Trends Across Channels")
    
```










### Payment Delivery Trends
There is an Upward Trend for **Payment Delivery** 
```{r}
ecol_pay_tbl %>% 
    mutate(pymt_date = floor_date(Pmt_dt, "year") %>% ymd()) %>% 
    group_by(pymt_date, Dlvry_status) %>% 
    summarise(Amount = sum(Amount)) %>% 
    
    ungroup() %>% 
    
    ggplot(aes(pymt_date, Amount)) +
    geom_point() +
    geom_line(size = 0.5, linetype = 1) +
    geom_smooth(span = 0.2, method = "loess") +
    facet_wrap(~Dlvry_status) +
    expand_limits(y = 0) +
    scale_y_continuous(labels = scales::comma) +
    theme_tq() +
    scale_fill_tq() +
    labs(x = "Year",
         y = "Payment Amount",
         title = "Payment Delivery Trends")
```











```{r}
ecol_pay_tbl %>% 
    mutate(pymt_date = floor_date(Pmt_dt, "year") %>% ymd()) %>% 
    group_by(pymt_date, State) %>% 
    summarise(Amount = sum(Amount)) %>% 
    
    ungroup() %>% 
    
    ggplot(aes(pymt_date, Amount)) +
    geom_point() +
    geom_line(size = 0.5, linetype = 1) +
    geom_smooth(span = 0.2, method = "loess") +
    facet_wrap(~State, ncol = 3) +
    expand_limits(y = 0) +
    scale_y_continuous(labels = scales::comma) +
    theme_tq() +
    scale_fill_tq() +
    labs(x = "Year",
         y = "Payment",
         title = "Payment State are Mostly Reconciliation Payment and its Upward")

```










### 2018 & 2019 Channel Comparison
**IVR and POS Channels** seems not to be **growing Month by Month** Compared to 2018 Payments Channels Transactions.
**CORP Channel is the Best Growth Channel 2019** than any other Channel.
```{r}
pct_ecol_tbl <- ecol_pay_tbl %>% 
    filter(!is.na(Pmtmthd_pmt_channel)) %>% 
    count(Pmtmthd_pmt_channel, State) %>% 
    group_by(Pmtmthd_pmt_channel, State, n) %>% 
    summarise(quant = sum(n)) %>% 
    
    ungroup() %>% 
    
    group_by(Pmtmthd_pmt_channel) %>% 
    mutate(pct = quant/sum(quant)) %>% 
    
    ungroup() %>% 
    
    mutate(Paymt_channel = as.factor(Pmtmthd_pmt_channel) %>% fct_rev()) %>% 
    mutate(cust_num = as.numeric(Paymt_channel))

mth_ecol_tbl <- ecol_pay_tbl %>% 
    filter(!is.na(Pmtmthd_pmt_channel)) %>%
    mutate(Month = month(Pmt_dt, label = TRUE, abbr = TRUE)) %>% 
    count(Pmtmthd_pmt_channel, Month) %>% 
    group_by(Pmtmthd_pmt_channel, Month, n) %>% 
    summarise(quant = sum(n)) %>% 
    
    ungroup() %>% 
    
    group_by(Pmtmthd_pmt_channel) %>% 
    mutate(pct = quant/sum(quant)) %>% 
    
    ungroup() %>% 
    
    mutate(Paymt_channel = as.factor(Pmtmthd_pmt_channel) %>% fct_rev()) %>% 
    mutate(cust_num = as.numeric(Paymt_channel))


Yr_tbl <- ecol_pay_tbl %>% 
    filter(!is.na(Pmtmthd_pmt_channel)) %>%
    mutate(Year = year(Pmt_dt)) %>% 
    mutate(Month = month(Pmt_dt, label = TRUE, abbr = TRUE)) %>% 
    count(Pmtmthd_pmt_channel, Year, Month) %>% 
    group_by(Pmtmthd_pmt_channel, Year, Month, n) %>% 
    summarise(quant = sum(n)) %>% 
    
    ungroup() %>% 
    
    group_by(Pmtmthd_pmt_channel) %>% 
    mutate(pct = quant/sum(quant)) %>% 
    
    ungroup() %>% 
    
    mutate(Paymt_channel = as.factor(Pmtmthd_pmt_channel) %>% fct_rev()) %>% 
    mutate(cust_num = as.numeric(Paymt_channel))

Yr_tbl %>% 
    filter(Year == 2018 | Year == 2019) %>% 
    ggplot(aes(Paymt_channel, Month)) +
    geom_tile(aes(fill = pct)) +
    geom_text(aes(label =scales::percent(pct)), size = 3) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "none") +
    scale_fill_gradient(low = "white", high = palette_light()[1]) +
    facet_wrap(~Year) +
    labs(x = "Payment Channel",
         y = "Month",
         title = "Payment Channel Comparison for 2018 and 2019")

```











### Payment Channel Preference for Entities
The Payment Channel of **Preference** for the **Leading** Payment Entity is **Corporate Payment** (Ministry of Mineral Resources) and for **Ministry of Finance** the leading Payment channels Preference are **POS, Bank Teller and CAM**
```{r}

# Heatmap of Entity Trend Dataframe
partner_ecol_tbl <- ecol_pay_tbl %>% # Create a Dataframe of Partner trend
    filter(!is.na(Pmtmthd_pmt_channel)) %>% # Remove NA's
    count(Pmtmthd_pmt_channel, Partner_code) %>% # Create count of Channel and Entity
    group_by(Pmtmthd_pmt_channel, Partner_code, n) %>% 
    summarise(quant = sum(n)) %>% # Sum the count column (n)
    ungroup() %>% 
    group_by(Pmtmthd_pmt_channel) %>% 
    mutate(pct = quant/sum(quant)) %>% # Percent Proportion of Entity Count
    ungroup() %>% 
    mutate(Paymt_channel = as.factor(Pmtmthd_pmt_channel) %>% fct_rev()) %>% 
    mutate(cust_num = as.numeric(Paymt_channel))



partner_ecol_tbl %>% 
    ggplot(aes(Paymt_channel, Partner_code)) +
    geom_tile(aes(fill = pct)) +
    geom_text(aes(label = scales::percent(pct)), size = 3) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "none") +
    scale_fill_gradient(low = "white", high = palette_light()[1]) +
    labs(x = "Payment Channel",
         y = "Entity")
```











### Partner Payment Preference
**MOF** accounts for all **POS** Channel Payments 
```{r}
partner_ecol_tbl <- ecol_pay_tbl %>% 
    filter(!is.na(Pmtmthd_pmt_channel)) %>% 
    count(Pmtmthd_pmt_channel, Partner_code) %>% 
    group_by(Pmtmthd_pmt_channel, Partner_code, n) %>% 
    summarise(quant = sum(n)) %>% 
    
    ungroup() %>% 
    
    group_by(Pmtmthd_pmt_channel) %>% 
    mutate(pct = quant/sum(quant)) %>% 
    
    ungroup() %>% 
    
    mutate(Paymt_channel = as.factor(Pmtmthd_pmt_channel) %>% fct_rev()) %>% 
    mutate(cust_num = as.numeric(Paymt_channel))



partner_ecol_tbl %>% 
    ggplot(aes(Paymt_channel, Partner_code)) +
    geom_tile(aes(fill = pct)) +
    geom_text(aes(label = scales::percent(pct)), size = 3) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "none") +
    scale_fill_gradient(low = "white", high = palette_light()[1]) +
    labs(x = "Payment Channel",
         y = "Entity")
```











### Percentage Yearly Growth
In **2010** There was a **significante** growth for Payment about **68790%**. Therefore **2010 Payment pattern** needs to be replicated to increase Growth for the coming years
```{r}
Yr_col_tbl <- ecol_pay_tbl %>% 
    mutate(Year = year(Pmt_dt)) %>% 
    group_by(Year) %>% 
    summarise(Amount = sum(Amount))


# mth cal 2019
mth_col_tbl <- ecol_pay_tbl %>% 
    mutate(Month = month(Pmt_dt, label = TRUE, abbr = FALSE)) %>% 
    mutate(Year = year(Pmt_dt)) %>% 
    filter(Year == 2019) %>% 
    group_by(Month) %>% 
    summarise(Amount = sum(Amount))


# mth cal 2018
mth_col_2018_tbl <- ecol_pay_tbl %>% 
    mutate(Month = month(Pmt_dt, label = TRUE, abbr = FALSE)) %>% 
    mutate(Year = year(Pmt_dt)) %>% 
    filter(Year == 2018) %>% 
    group_by(Month) %>% 
    summarise(Amount = sum(Amount))


Yr_col_tbl %>% 
    mutate(Amount_lag = lag(Amount, n = 1)) %>% 
    mutate(Amount_lag = case_when(
        is.na(Amount_lag) ~ Amount,
        TRUE ~ Amount_lag
    )) %>% 
    
    mutate(diff_1 = Amount - Amount_lag) %>% 
    mutate(pct_diff = diff_1 / Amount_lag) %>% 
    mutate(pct_diff_char = scales::percent(pct_diff, accuracy = 1)) %>% 
    select(Year, Amount, pct_diff_char) %>% 
    rename(Percent = pct_diff_char)
    
```












### 2018 & 2019 Monthly Trend
There is similar **Quarterly Trends** for 2018 and 2019.
**Q2 2019** was far better than **Q2 2018**, However **July 2018 Outperform** July 2019 and there is a noticeable **downward sharp trend** for **August 2018 and 2019**.
```{r}
calculate_pct_diff <- function(data) {
    data %>% 
        mutate(Amount_lag = lag(Amount, n = 1)) %>% 
        mutate(Amount_lag = case_when(
            is.na(Amount_lag) ~ Amount,
            TRUE ~ Amount_lag
        )) %>% 
        
        mutate(diff_1 = Amount - Amount_lag) %>% 
        mutate(pct_diff = diff_1 / Amount_lag) %>% 
        mutate(pct_diff_char = scales::percent(pct_diff, accuracy = 1)) %>% 
        select(Month, Amount, pct_diff_char) %>% 
        rename(Percent = pct_diff_char)
    
}


m_2019 <- mth_col_tbl %>% 
    calculate_pct_diff()


m_2018 <- mth_col_2018_tbl %>% 
    calculate_pct_diff()

trend_data <- m_2019 %>% 
    bind_cols(m_2018 %>% filter(!Month %in% c("September", "October", "November", "December"))) %>% 
    rename(Pct_2019 = Percent) %>% 
    rename(Pct_2018 = Percent1) %>% 
    select(-Month1) %>% 
    rename(Amount_2019 = Amount) %>% 
    rename(Amount_2018 = Amount1)



trend_data %>% 
    select(Month, Amount_2018, Amount_2019, everything()) %>% 
    gather(key = "Year_Amount", value = "Amount", -Pct_2019, -Pct_2018, -Month) %>% 
    group_by(Month, Year_Amount) %>% 
    summarise(Amount = sum(Amount)) %>% 
    ungroup() %>% 
    ggplot(aes(x = Month, y = Amount, group = Year_Amount)) +
    geom_point() +
    geom_line(aes(color = Year_Amount)) +
    #geom_density() +
    #geom_line(size = 0.5, type = "l") +
    geom_smooth(span = 0.2, method = "loess") +
    expand_limits(y = 0) +
    scale_y_continuous(labels = scales::comma) +
    theme_tq() +
    scale_fill_tq() +
    labs(x = "Month",
         y = "Payment Amount",
         title = "Monthly Trend for 2018 and 2019")

```










```{r}
umap_kmeans_3_result_tbl %>% # using the merged umap and clustering table
    distinct(Partner_code, .keep_all = TRUE) %>% # select the distinct entity
    mutate(label_text = str_glue("Entity: {Partner_code} 
                                 Cluster: {.cluster}")) %>% # create label_text to display entity and the assigned cluster
    ggplot(aes(X, Y, color = .cluster)) +
    geom_point() +
    geom_label_repel(aes(label = label_text), size = 3) +
    theme_tq() +
    scale_color_tq() +
    labs(
        title = "eCollection Payment Entity Segmentation",
        subtitle = "UMAP Kmeans Clustering"
    ) +
  theme(legend.position = "none")
```











```{r}
cluster_label_tbl <- tibble(
  .cluster = 1:3,
  .cluster_label = c(
    "Internet Channel, Low Payment, Paid, BLN/G2G",
    "BTELLER/Internet, Low Payment, Paid/Recon, BLN",
    "INTERNET/COROP/BTELLER, High/Medium/Low, Mostly G2G, RECON"
  )
) %>% 
  mutate(.cluster = as_factor(as.character(.cluster)))



umap_kmeans_3_result_tbl %>% 
  left_join(cluster_label_tbl) %>% 
  distinct(Partner_code, .keep_all = TRUE) %>% 
  mutate(label_text = str_glue("Entity: {Partner_code}
                                 Cluster: {.cluster}
                               {.cluster_label}")) %>% 
  ggplot(aes(X, Y, color = .cluster)) +
  geom_point() +
  geom_label_repel(aes(label = label_text), size = 3) +
  theme_tq() +
  scale_color_tq() +
  labs(x = "",
       y = "",
    title = "eCollection Payment Entity Segmentation 2D Projection",
    subtitle = "UMAP Kmeans Clustering"
  ) +
  theme(legend.position = "none")

```


There are **3 distinct categories** of Entities with respect to Payment Channel Preference, Payment amount size, Payment Status and Source of Channels for Payments.

* Cluster 1 Entities make most payments with Internet Channel and are mostly low payments amount (payments less than SAR 16,081.24) with Payment source of BLN/G2G source.

* Cluster 2 Entities make payments with Bank Teller and Internet Channels, mostly low payment amounr with Status of payment, split between Reconciliation and Paid status, with BLN source channel.

* Cluster 3 Entities make payments with Internet/Corp and Bank Teller channels for High and Medium Payments amount (>= 731870.91) and mostly with G2G source channels.


